\input{formattingHeader}

\title{Checking the OpenLCB Datagram Transport Protocol}

\begin{document}
\maketitle

\introductionCaveats
{https://nbviewer.org/github/openlcb/documents/blob/master/standards/DatagramTransportS.pdf}
{Datagram Transport Standard}.

\section{Datagram Transport Procedure}

\checkProcedure{Datagram Checking}

Note that this process is unable to check datagrams from the DBC to the checker.
There is no standard way to elicit those.  They will be checked as part of the 
Memory Configuration Protocol checking, if applicable.

A node which does not self-identify in PIP that it supports
datagram transfer will be considered to have passed these checks.
\pipsetFootnote

\subsection{Datagram Reception}

This checks the messages in Standard sections 4.1 and 4.2 or 4.3, 
including the error code constraints in Standard section 4.3.1
\footnote{This section references the Message Transport Standard, 
section 3.5.5}
and the interactions in sections 6.1 or 6.2.

The checker will send a datagram to the DBC. The DBC in turn will 
either accept the datagram (sections 4.2 and 6.1) or 
reject it (sections 4.3 and 6.2). Either response is acceptable.

This section's checks cover:

\begin{enumerate}
\item If the datagram is accepted,
    \begin{enumerate}
    \item That the Datagram OK message is received,
    \item That the Datagram OK message has proper source and destination node IDs, 
    \item that the Datagram OK message contains exactly 1 byte of flags,
    \item that the 0x70 bits of the flags are zeros.
    \end{enumerate}
\item if the datagram is rejected, 
    \begin{enumerate}
    \item That the Datagram Rejected message is received,
    \item That the Datagram Rejected message has proper source and destination node IDs, 
    \item that the Datagram Rejected message contains at least two bytes of error code,
    \item that the 0xF000 bits of the error code are either 0x1000 or 0x2000,
    \item that the 0x0F00 bits of the error code are zeros.
    \end{enumerate}
\end{enumerate}

This process is repeated for datagrams of length 1, 10 and 72 bytes to exercise the full
range of datagram framing.


\end{document}  
