\input{formattingHeader}

\titleandheader{Checking the OpenLCB Function Description Information Standard}

\begin{document}
\maketitle
\thispagestyle{firststyle}

\introductionCaveats
    {https://nbviewer.org/github/openlcb/documents/blob/master/standards/FunctionDescriptionInformationS.pdf}
    {Function Definition Information Standard}

\section{Function Description Information Procedure}

\checkProcedure{Function Description Information Checking}

This plan assumes that the Datagram Transport Protocol and the Memory Configuration Protocol 
have been separately checked. It uses those, but does not do any detailed checking of them.

This procedure checks the operation of a specific Train Node. 
The Node ID of the node to be checked needs to be specified; 
if there are previous steps necessary to e.g. create a specific 
virtual train node, those should be executed before the checks in each section.

\subsection{FDI Memory Present checking}

This section checks that the FDI memory defined in the Function 
Definition Information Standard is present.

The Get Address Space Information Command from the Memory Configuration Protocol
is used to validate that space 0xFA is present and read-only.

\subsection{Validation checking}

This section checks the content of the FDI against the XML Schema 
defined in Section 5 of the Standard to make sure that it is valid XML. 

This check reads the information from the 0xFA memory space
until it encounters a null character.
\begin{enumerate}
\item If the space is not available, or there are no bytes of content in it, the check fails.
\item If the read sequence terminates early, without ending in a null character, the check fails.
\end{enumerate}
 
The check then compares the length of the 0xFA memory space 
as specified by the Memory Configuration Protocol Standard sections 4.15 and 4.16
vs. the actual length of the retrieved FDI.
If they do not agree, the check fails.

The check then compares the first two lines of the content against the 
definition in section 5. 
The location of the schema in the xsi:noNamespaceSchemaLocation attribute 
must match the Standard.
Any valid version number for the FDI schema is accepted.

The check then validates the content against the FDI XML Schema
retrieved from the given FDI schema location.

\end{document}  
