\input{formattingHeader}

\titleandheader{Checking the OpenLCB Firmware Upgrade Protocol Standard}

\begin{document}
\maketitle
\thispagestyle{firststyle}

\introductionCaveats
    {https://nbviewer.org/github/openlcb/documents/blob/master/standards/FirmwareUpgradeS.pdf}
    {Firmware Upgrade Protocol Standard}

\section{Firmware Upgrade Protocol Procedure}

\checkProcedure{Firmware Upgrade Protocol Checking}

A node which does not self-identify in PIP that it supports
the Firmware Upgrade Protocol should be considered to have passed these checks.
\pipsetFootnote

This check requires a vendor-provided current version of the 
appropriate firmware file for downloading into the node being checked.
Do not start this check without having that file available,
as a node is not required to recover from a firmware upgrade attempt without
being successfully upgraded.

\subsection{Firmware Upgrade status}

This check follows the sequence documented in Section 5.5

At this time, only the datagram-based transfers are checked.

\begin{enumerate}
\item The checked uses SNIP to retrieve the software version and displays that for manual 
    validation.
\item The checker sends a Memory Configuration datagram command “Freeze” with an 
    argument of Firmware Space 0xEF. 
    A Datagram Received reply may or many not be received in response.
\item The checker waits up to 20 seconds for a Node Initialization Complete message from the 
    node being checked.
\item The checker confirms that the PIP Firmware Upgrade active bit is set.
\item The checker uses Memory Write operations of 64 bytes to download the contents of the firmware file.
    to memory space 0xEF.
\item The checker resets the node being checked using a Memory Configuration 
    datagram command “Unfreeze” with an argument of Firmware Space 0xEF.    
    A Datagram Received reply may or many not be received in response.
\item The checker waits up to 20 seconds for a Node Initialization Complete message from the 
    node being checked.
\item The checker confirms that the PIP Firmware Upgrade active bit is reset.
\item The checker uses SNIP to retrieve the software version and displays that for manual 
    validation.
\end{enumerate}

\end{document}  
